---
title: "Solutions"
output:
  html_document:
    css: 'cayman.css'
date: '`r Sys.Date()`'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solutions  

Below are the solutions for the exercises. Many ways to complete these exercises... so don't take this as the only way.  

### Exercise #1  


```{r}
library(palmerpenguins)
data("penguins")

descriptive_stats_function <- function(data, cols) {
  
  cat(paste0('There are ', length(data), ' columns and ' , nrow(data), ' rows in this data.frame'))
  
  mean_data <- mean(data[,cols][[1]], na.rm = TRUE)
  sd_data <- sd(data[,cols][[1]], na.rm = TRUE)
  col_name <- cols[[1]]
  
  return(data.frame(mean = mean_data,
                    sd = sd_data,
                    name = col_name))
  
}

test_df <- descriptive_stats_function(penguins, cols = 'bill_length_mm')
test_df

descriptive_stats_function(penguins, cols = 'bill_depth_mm')
```


### Exercise #2  

Now we'll want to add another argument like a function. Again, lots of ways to do this but we'll just add a `vector` of function names that only take a vector



```{r}
descriptive_stats_function <- function(data, cols, function_name = c('mean', 'sd'), ...) {
  
  function_output <- list()
  
  for(i in function_name){
    
    call_function <- do.call(i, list(data[,cols][[1]], ...))
    
    call_df <- data.frame(i = call_function)
    
    colnames(call_df)[1] <- i
    
    function_output <- append(function_output, call_df)
    
  }
  
  return(rbind(function_output))
  
}

descriptive_stats_function(penguins, 'bill_length_mm', function_name = c('mean', 'sd', 'median', 'max', 'min'), na.rm = TRUE)

```


### Exercise #3  

Here we are trying to catch an error in our `function_name` argument.  


```{r}
descriptive_stats_function <- function(data, cols, function_name = c('mean', 'sd'), ...) {
  
  function_output <- list()
  
  for(i in function_name){
    
    call_function <- tryCatch(do.call(i, list(data[,cols][[1]], ...)),
                              error = function(e) {
                                NA_real_
                              })
    
    call_df <- data.frame(i = call_function)
    
    if(is.na(call_function)){
      
    colnames(call_df)[1] <- 'error'  
    
    } else {
      
    colnames(call_df)[1] <- i
    
    }
    
    function_output <- append(function_output, call_df)
    
  }
  
  return(rbind(function_output))
  
}

descriptive_stats_function(penguins, 'bill_length_mm', function_name = c('mean', 'median', function(x) x/100, 'hello', 'sd'), na.rm = TRUE)

```
